---
title: "Preliminary Exploration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{preliminary_exploration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = F, warning = F}
#library(RiskPredictionSims)
devtools::load_all()
library(lme4)
library(tidyverse)
library(mnormt)
```

The `create_data` function creates outcomes from the following model:

$$
\begin{aligned}
\Pr(Y_{ijt}=1) = \left(1+\exp\{-(\alpha + \beta x_{ijt} + \gamma_j + \omega_{jt})\}\right)^{-1}
\end{aligned}
$$

where $t=1,\ldots, T$ indexes time, $j=1,\ldots, J$ indexes centers, and $i=1,\ldots,n_{jt}$ indexes
runs, i.e. observations, within a center in a given year. The generating mechanism
has the following features:

* $T$ and $J$ are specified by `num_years` and `max_num_centers`, respectively
* $\alpha$ and $\beta$ are specified by `alpha` and `beta`, respectively
* $x_{ijt} \sim N(0,1)$ independently for $i$, $j$, and $t$
* $\gamma_j\sim N(0, \tau^2_\gamma)$ independently for each $j$, where $\tau^2_\gamma$ is
specified by `gamma_var` 
* $\{\omega_{j1},\omega_{j2},\ldots\} \sim MVN(0, \tau^2_\omega \Sigma_\omega)$, 
independently for each $j$, where $\tau^2_\omega$ is specified by `omega_var` and
$\Sigma_\omega$ is an AR(1) matrix with 1s along the diagonal and correlation parameter
specified by `omega_cor`
* $n_{jt}|\lambda_j \sim \textrm{Pois}(\lambda_j)$ independently (conditional on $\lambda_j$)
for each $j$ and $t$. $\lambda_j$ can be specified by `mean_runs_per_center` or
if not provided it will be drawn from a shifted-folded-$t$ distribution, independently
for each $j$, to create skewness in the number of runs per center

Below we create a dataset using its default values (see function for default values). 
```{r}
data1 <- create_data()
```

Fit a generalized linear mixed effects models with year-center random effects. 
```{r}
mod_data1 <-
  glmer(y ~ x + (1|year_id:center_id), family = "binomial", data = data1)
```
Now extract the BLUPS and join with the simulated data to facilitate comparisons.

```{r}
summary_mod_data1 <- 
  ranef(mod_data1)$`year_id:center_id` %>% 
  rownames_to_column() %>%
  separate(rowname,into = c("year_id", "center_id")) %>%
  rename("estimated_raneff" = "(Intercept)") %>%
  mutate(year_id = as.integer(year_id), 
         center_id = as.integer(center_id)) %>%
  left_join(data1 %>% 
              group_by(center_id, year_id) %>% 
              summarize(num_runs = n(),
                        true_raneff = mean(true_raneff),
                        .groups = "drop"),
            by = c("year_id", "center_id")) %>%
  mutate(num_runs_cat = cut(num_runs, breaks = c(0, 1, 4, 20, Inf))) %>%
  group_by(num_runs_cat) %>%
  mutate(sp_cor = round(cor(true_raneff, estimated_raneff, method = "spearman"), 3)) %>%
  ungroup()
```

How do the estimated random effects compare to the true random effects for each
center and year:
```{r, message = FALSE, fig.width = 7, fig.height = 5}
ggplot(summary_mod_data1, 
       aes(x = true_raneff, y = estimated_raneff)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  geom_abline(slope = 1, intercept = 0) + 
  geom_text(data = distinct(summary_mod_data1, num_runs_cat, .keep_all = TRUE),
            aes(x = -1.5, y = 0.75, label = sp_cor)) + 
  facet_wrap(vars(num_runs_cat))

```

